Making use of spatially biased variables in ecosystem condition accounting – a GIS based workflow

### 2025-04-22 manuscript/Manuscript.qmd

library(tidyverse)
library(knitr)
library(sf)
library(tmap)
library(tmaptools)
library(stars)
library(terra)
library(tidyterra)
library(ggtext)
library(cowplot)
library(units)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggmagnify)
library(ggridges)
library(eaTools) #https://ninanor.github.io/eaTools/ version 0.0.0.9000
library(ggpubr)
library(kableExtra)

@@

# field survey
# # downloaded from https://kartkatalog.geonorge.no/metadata/naturtyper-miljoedirektoratets-instruks/eb48dd19-03da-41e1-afd9-7ebc3079265c
path_naturetypes <- "../data/survey.gdb"

# municipality outline
path_muni <- "../data/Basisdata_0000_Norge_25833_Kommuner_FGDB.gdb"

@@

# I already did some work to identify the relevant nature types
# summary file (https://github.com/NINAnor/ecosystemCondition/blob/main/data/naturetypes/natureType_summary.rds)
naturetypes_summary <- readRDS("../data/natureType_summary.rds")

# Survey data
# The data is too big to be stored on GitHub
# Import polygon data set
st_layers(path_naturetypes)
naturetypes <- sf::st_read(dsn = path_naturetypes, layer = "naturtyper_nin_omr")
# 142k polygons (2023)

@@

# Outline of norway (coastline)
outline <- sf::read_sf("../data/outlineOfNorway_EPSG25833.shp") |>
  st_transform(myCRS)

@@

# subset of the three target municipalities
muni3 <- muni |>
  filter(kommunenummer %in% c(
    "3020", # Nordre Follo
    "3451", # Nord-Aurdal
    "3446" # Gran
  )) |>
  mutate(Municipality = case_when(
    kommunenummer == "3020" ~ "Nordre Follo",
    kommunenummer == "3451" ~ "Nord-Aurdal",
    kommunenummer == "3446" ~ "Gran"
  ))

# To crop EDM, I need the three municipalities seprately.
nf <- muni3 |>
  filter(kommunenummer == "3020")
na <- muni3 |>
  filter(kommunenummer == "3451")
gr <- muni3 |>
  filter(kommunenummer == "3446")
```

@@

We wish to thank Kwaku Peprah Adjei for statistical help, and Hanno Sandvik for code review.

>>>>

library(tidyverse)
library(knitr)
library(sf)
library(tmap)
library(tmaptools)
library(stars)
library(terra)
library(tidyterra)
library(ggtext)
library(cowplot)
library(units)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggmagnify)
library(ggridges)
library(eaTools) #https://ninanor.github.io/eaTools/ version 0.0.0.9000
library(ggpubr)
library(kableExtra)
library(here)

@@

# field survey
# # downloaded from https://kartkatalog.geonorge.no/metadata/naturtyper-miljoedirektoratets-instruks/eb48dd19-03da-41e1-afd9-7ebc3079265c
path_naturetypes <- "P:/41001581_egenutvikling_anders_kolstad/data/Naturtyper_nin_0000_norge_25833_FILEGDB.gdb"

# municipality outline
path_muni <- here("data/Basisdata_0000_Norge_25833_Kommuner_FGDB.gdb")

@@

# I already did some work to identify the relevant nature types
# summary file (https://github.com/NINAnor/ecosystemCondition/blob/main/data/naturetypes/natureType_summary.rds)
naturetypes_summary <- readRDS(here("data/natureType_summary.rds"))

# Survey data
# The data is too big to be stored on GitHub
# Import polygon data set
st_layers(path_naturetypes)
Sys.setlocale("LC_ALL","Norwegian")
naturetypes <- sf::st_read(dsn = path_naturetypes, layer = "naturtyper_nin_omr")
# 142k polygons (2023)

@@

# Outline of norway (coastline)
outline <- sf::read_sf(here("data/outlineOfNorway_EPSG25833.shp")) |>
  st_transform(myCRS)

@@

# subset of the three target municipalities
muni3 <- muni |>
  filter(kommunenummer %in% c(
    "3207", # Nordre Follo
    "3451", # Nord-Aurdal
    "3446" # Gran
  )) |>
  mutate(Municipality = case_when(
    kommunenummer == "3207" ~ "Nordre Follo",
    kommunenummer == "3451" ~ "Nord-Aurdal",
    kommunenummer == "3446" ~ "Gran"
  ))

# To crop EDM, I need the three municipalities seprately.
nf <- muni3 |>
  filter(kommunenummer == "3207")
na <- muni3 |>
  filter(kommunenummer == "3451")
gr <- muni3 |>
  filter(kommunenummer == "3446")
```

@@

We wish to thank Kwaku Peprah Adjei for statistical help, and Hanno Sandvik for code review.
We acknowledge partial funding from the Research Council of Norway project EcoGaps (contract #320042).
